---
title: "decathlon_anlysis"
output: html_document
---

1) The contestant with the longest long jump seen in the data?

```{r}

longest_long_jump <- decathlon_clean %>%
  select(name, long_jump) %>% 
  slice_max(long_jump, n = 1)

longest_long_jump
   
```

2) The average 100m time in each competition

```{r}

decathlon_clean %>% 
  select(x100m) %>% 
  summarise("average 100m" = mean(x100m))

```

3) The contestant with the highest total points across both competitions, i.e the Decaster and OlympicG.

```{r}
decathlon_clean %>%
  select(name, points,competition) %>% 
  filter(competition == "Decastar") %>% 
  arrange(desc(points)) %>% 
  slice_max(1)

decathlon_clean %>%
  select(name, points,competition) %>% 
  filter(competition == "OlympicG") %>% 
  arrange(desc(points)) %>% 
  slice_max(1)

```

4) The shot-put scores for the top three competitors in each competition.

```{r}

decathlon_clean %>%
  select(name, shot_put,competition) %>% 
  filter(competition == "Decastar") %>% 
  slice_max(shot_put, n = 3)

decathlon_clean %>%
  select(name, shot_put, competition) %>% 
  filter(competition == "OlympicG") %>% 
  slice_max(shot_put, n = 3)


```

5) Average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds.

```{r}
  
avg_below_50s <- decathlon_clean %>% 
  select(x400m) %>% 
  filter(x400m < 50) %>% 
  summarise(mean(x400m)) %>% 
  pull()

avg_above_50s <- decathlon_clean %>% 
  select(x400m) %>% 
  filter(x400m > 50) %>% 
  summarise(mean(x400m)) %>% 
  pull()

comparison <- list("avg_below_50s" = avg_below_50s, 
                   "avg_above_50s" = avg_above_50s)

comparison

```